---
output: 
  word_document:
    reference_docx: C:\Users\Martin Pastoors\AppData\Roaming\Microsoft\QuickStyles\Rmarkdown style.dotx
---

```{r setup, include=FALSE}

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	crop = TRUE,	comment = "")
knit_hooks$set(crop = hook_pdfcrop)

opts_knit$set(root.dir="D:/SPRFMO/2017")

```


```{r initialization, include=FALSE}
# ---------------------------------------------------------------------------------------
# EU Jack mackerel CPUE analysis in SPRFMO area 
#
# 23/09/2017 First coding of the CPUE analysis
# ---------------------------------------------------------------------------------------

rm(list=ls())

# Libraries
library(rmarkdown)     # rmarkdown functionality
library(pander)        # tables
library(lubridate)     # data handling
library(reshape2)      # reshaping data; e.g. cast
library(readxl)        # excel reader
library(broom)         # clean up statistics
library(scales)        # pretty scales
library(stringr)       # string manipulations
library(tidyverse)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble

# default settings for tables
panderOptions('table.alignment.default', function(df) ifelse(sapply(df, is.numeric), 'right', 'left'))

# Load map data
load("D:/XXX/prf/rdata/world.df.RData")
load("D:/XXX/prf/rdata/eez.df.RData")
load("D:/XXX/prf/rdata/fao.df.RData")

# Source all the utilities
source("D:/GIT/mptools/R/my_utils.r")

#  Load data
load(file="rdata/eu_cjm.RData")

```

## EU Fisheries CPUE analysis for Jack mackerel in the SPRFMO area

### Subtitle

&nbsp;  

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`


##### page break

### 1. Introduction

The assessment of Jack Mackerel in the southern Pacific is based on many different sources of information, including the nominal Catch per Unit Effort (expressed as catch per day) of the EU fleet. The use of nominal CPUE for calibrating stock assessments is known to be potentially problematic and therefore SPRFMO (2011) recommended that to serve as indices of abundance, the CPUE should be standardized to take into account factors such as historical changes in vessels, fishing areas, seasonal fishing patterns and environmental
factors. This standardization approach has already been applied by China (Li et al, 2013). 

In this document, the catch and effort data for the EU fleet is analysed with the aim to develop a standardized CPUE series. 

[ more text ]

### 2. Material and methods

[ Overview of all data available (tables, figures) ]

```{r echo=FALSE, fig.width=10, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

cjm %>% 
  rbind.all.columns(data.frame(year = 2012)) %>% 
  
  ggplot(aes(x=shootlon, y=shootlat)) +
  theme_publication() +
  theme(
    axis.title = element_blank(), 
    legend.position = "none") +
  coord_quickmap(xlim=c(-120,-50) , ylim=c(-50,-10)) +
  geom_polygon(data=fao.df, aes(long, lat, group=group), fill = NA, size=0.25, color="gray60", alpha=0.3) +
  geom_polygon(data=world.df, aes(long,lat,group=group), fill="gray30", size=0.25,color="gray15", alpha=0.7) +
  geom_jitter(aes(colour=vesselname), size=0.8, alpha=0.5) +
  ggtitle("Haul positions (averaged by day")
  facet_wrap(~year, drop=FALSE, ncol=4)

```

*Table 1. Summary of EU catch and effort data.*


```{r echo=FALSE, fig.width=10, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

cjm %>% 
  rbind.all.columns(data.frame(year = 2012)) %>% 
  
  ggplot(aes(x=shootlon, y=shootlat)) +
  theme_publication() +
  theme(
    axis.title = element_blank(), 
    legend.position = "none") +
  coord_quickmap(xlim=c(-120,-50) , ylim=c(-50,-10)) +
  geom_polygon(data=fao.df, aes(long, lat, group=group), fill = NA, size=0.25, color="gray60", alpha=0.3) +
  geom_polygon(data=world.df, aes(long,lat,group=group), fill="gray30", size=0.25,color="gray15", alpha=0.7) +
  geom_jitter(aes(colour=vesselname), size=0.8, alpha=0.5) +
  ggtitle("Haul positions (averaged by day")
  facet_wrap(~year, drop=FALSE, ncol=4)

```

*Figure 1. Haul positions (averaged by day) 2007-2017). Colours indicate the different EU vessels participating.*

lajflaj

```{r echo=FALSE, fig.width=10, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

t <- 
  cjm %>% 
  rbind.all.columns(data.frame(year = 2012)) %>% 
  filter(!is.na(shootlon)) %>% 
  filter(!is.na(shootlat)) %>% 
  mutate(rect = encode_zchords(x=shootlon, y=shootlat, dx = 1, dy = 0.5) ) %>% 

  group_by(vesselname, year, day, rect) %>% 
  summarise(catch  = sum(catch, na.rm=TRUE), 
            catch2 = sum(catch2, na.rm=TRUE)) %>% 
  group_by(year, rect) %>% 
  summarise(catch  = mean(catch, na.rm=TRUE), 
            catch2 = mean(catch2, na.rm=TRUE)) %>% 
  separate(rect, c("shootlon", "shootlat"), sep = ":", convert = TRUE, remove = FALSE) %>% 
  
  ungroup()

b <- 
  log_breaks(n=7)(c(1,max(select(t, catch), na.rm=TRUE))) 

td <-
  t %>% 
  mutate(catch = cut(catch,breaks=b, include.lowest=T, dig.lab=10) ) %>% 
  mutate(catch2= cut(catch2,breaks=b, include.lowest=T, dig.lab=10) ) %>% 
  filter(!is.na(catch))

td %>% 
  ggplot(aes(x=shootlon, y=shootlat)) +
  theme_publication() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        text             = element_text(size=12),
        legend.key.width  = unit(1, "cm") ) +  
  
  coord_quickmap(xlim=c(-120,-50) , ylim=c(-50,-10)) +
  geom_polygon(data=fao.df, aes(long, lat, group=group), fill = NA, size=0.25, color="gray60", alpha=0.3) +
  geom_polygon(data=world.df, aes(long,lat,group=group), fill="gray30", size=0.25,color="gray15", alpha=0.7) +
  geom_tile(aes(shootlon, shootlat, fill = catch), colour=NA, alpha=1.0) +
  scale_fill_brewer(palette = "YlOrRd") + 
  labs(x = NULL, y = NULL) +
  ggtitle("Haul positions (averaged by day") +
  facet_wrap(~year, drop=FALSE, ncol=4)

```


*Figure 2. Catch of Jack mackerel (summed by 1 degree longitude and 0.5 degree latitude) in 2007-2017. Catch in tonnes expressed on a log scale.*

```{r echo=FALSE, fig.width=10, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

m <-
  cjm %>% 
  # filter(year != 2012) %>% 
  group_by(year) %>%
  summarise(lcatch2 = mean(lcatch2, na.rm=TRUE), 
            day     = mean(day, na.rm=TRUE)) %>% 
  mutate(date = as.numeric(year) + day/365) %>% 
  data.frame() %>% 
  rbind.all.columns (data.frame(year=2012))
  
cjm %>% 
  mutate(date = year + day/365) %>% 
  ggplot(aes(date, lcatch2)) +
  theme_publication() +
  theme(legend.position="none")+
  geom_jitter(aes(colour=vesselcode)) +
  geom_line(data=m, aes(date, lcatch2), colour="gray20", linetype="dashed", size=1) +
  geom_point(data=m, aes(date, lcatch2)) +
  scale_x_continuous(breaks=seq(2005, 2017, by = 5)) 

```

*Figure 3. Jack mackerel CPUE (log(catch / day)). Colours indicate the different vessels.*

```{r echo=FALSE, fig.width=10, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

m <-
  cjm %>% 
  filter(year != 2012) %>% 
  filter(lcatch2 >= 0.001) %>% 
  filter(shootlat2 %in% c(-20, -30, -40)) %>% 
  filter(!grepl("NA", area)) %>% 
  group_by(year, area, shootlat2, shootlon2) %>%
  summarise(lcatch2 = mean(lcatch, na.rm=TRUE), 
            day     = mean(day, na.rm=TRUE)) %>% 
  mutate(date = as.numeric(year) + day/365) 

cjm %>% 
  filter(!grepl("NA", area)) %>% 
  filter(lcatch2 >= 0.001) %>% 
  filter(shootlat2 %in% c(-20, -30, -40)) %>% 
  mutate(date = year + day/365) %>% 
  mutate(shootlat2 = factor(shootlat2, levels = c("-20", "-30", "-40"))) %>% 
  # group_by(shootlat2) %>% filter(row_number() ==1) %>% select(shootlat2) %>% View()

  ggplot(aes(date, lcatch2, group=area)) +
  theme_publication() +
  theme(legend.position="none") +
  geom_jitter(colour="lightblue", alpha=0.5) +
  geom_line(data=m, aes(x=date, y=lcatch2), colour="gray20", linetype="dashed", size=1, inherit.aes = FALSE) +
  scale_x_continuous(breaks = seq(2005, 2017, by = 5)) +
  facet_grid((shootlat2)~shootlon2)


```


### 3. Results

### 4. Discussion and conclusions

### 5. References

Li, G., X. Zou, X. Chen, Y. Zhou and M. Zhang (2013). "Standardization of CPUE for Chilean jack mackerel (Trachurus murphyi) from Chinese trawl fleets in the high seas of the Southeast Pacific Ocean." Journal of Ocean University of China 12(3): 441-451.

SPRFMO (2011) Report of the Jack Mackerel Subgroup. Tenth Science Working Group of SPRFMO, 19 – 23 September
2011, Port Vila, Vanuatu.

