---
output: 
  word_document:
    reference_docx:  ../PFA_report_template_v1.7.dotx
---

```{r setup, include=FALSE}

# ==================================================================
# SPRFMO Catch template reader (SC8 2020)
#
# 11/10/2019: first coding
# 03/10/2020: finalized coding for 2019 and 2020 data
# 04/10/2020: also available for earlier data now
# 11/08/2021: added the 2020 ALK data from Chile (both old and new method)
# ==================================================================

# Reset lists
# rm(list=ls())

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

# general libraries
library(readxl)        # excel reader from hadley; much quicker than the java based excel readers
library(lubridate)     # data and time functions
library(stringr)       # string manipulation
library(pander)        # for tables
library(reshape2)      # cast and melt dataframes
library(scales)        # pretty scales
library(tidyverse)     # data manipulation and piping

# source utilities
source("../../prf/r/my utils.R")

data_path <- "C:/DATA/Onedrive - PFA/Documents/SPRFMO/data submissions"


# Create list of directories
dir.list <- list.dirs(path=data_path, recursive=TRUE, full.names = TRUE)
dir.list <- dir.list[grepl("ALK",toupper(dir.list))]
dir.list <- dir.list[!grepl("ALK$",toupper(dir.list))]
dir.list <- dir.list[!grepl("prel$|old$",tolower(dir.list))]

# Create file list for import
file.list <- as.character(NA)
for (d in dir.list) {
  file.list   <- c(file.list,
                   list.files(path = d,recursive  = FALSE,pattern = "xls",full.names = TRUE,ignore.case= FALSE) )
}
file.list <- file.list[!grepl("\\~", file.list)]
# file.list <- file.list[!grepl("2017", file.list)]
file.list <- file.list[!is.na(file.list)]


# read_catch function -------------------------------------------------------------------
# i <- 5
# ws<-"catch data"

read_catch <- function(file, ws, member, species, year) {

  t1 <-
    suppressMessages(read_excel(file, sheet = ws, col_names = FALSE, na="NA", col_types = "text", range = "B12:M13")) %>%
    t() %>%
    data.frame(stringsAsFactors = FALSE) %>%
    setNames(c("quarter", "year")) %>%
    mutate(quarter = as.numeric(stringr::str_match_all(quarter, "[0-9]+")) ) %>%
    zoo::na.locf(.)                       # fill the empty values with previous values

  t <-
    suppressMessages(read_excel(file, sheet = ws, col_names = FALSE, na="NA", col_types = "text", range = "B15:M20", .name_repair = "minimal")) %>%
    t() %>%
    data.frame(stringsAsFactors = FALSE) %>%
    setNames(c("variable","unit","F1","F2","F3","F4")) %>%
    bind_cols(t1) %>%
    gather(key=fleet, value=value, 3:6) %>%
    filter(!is.na(value)) %>%
    mutate(unit = gsub("[[:punct:]]", "", tolower(unit))) %>%
    mutate(variable = tolower(variable)) %>%
    mutate(member=member, species=species, year=as.integer(year), file=basename(file))  %>%
    mutate(value = as.numeric(value))


  # generate dataframe is not all empty
  if (nrow(t) > 0) {
    print(ws)
    return(t)
  }

} # end of function

# i <- 1
# ws = "sampling"
# file = file.list[i]

# read_sampling function -------------------------------------------------------------------

read_sampling <- function(file, ws, member, species, year) {

  t1 <-
    suppressMessages(read_excel(file, sheet = ws, col_names = FALSE, na="NA", col_types = "text", range = "B8:Q8")) %>%
    t() %>%
    data.frame(stringsAsFactors = FALSE) %>%
    setNames(c("quarter")) %>%
    mutate(quarter = as.numeric(stringr::str_match_all(quarter, "[0-9]+")),
           quarter = ifelse(quarter > 4, NA, quarter),
           year    = year) %>%
    zoo::na.locf(.)                        # fill the empty values with previous values

  t <-
    read_excel(file, sheet = ws, col_names = FALSE, na="NA", col_types = "text", range = "B9:Q14", .name_repair = "minimal") %>%
    t() %>%
    data.frame(stringsAsFactors = FALSE) %>%
    setNames(c("variable","unit","F1","F2","F3","F4")) %>%
    bind_cols(t1) %>%
    gather(key=fleet, value=value, 3:6) %>%
    mutate(unit = gsub("[[:punct:]]", "", tolower(unit))) %>%
    mutate(member=member, species=species, year=year, file=basename(file))  %>%
    filter(!is.na(value)) %>%
    mutate(value = as.numeric(value))

  # generate dataframe is not all empty
  if (nrow(t) > 0) {
    print(ws)
    return(t)
  }

} # end of function

# read_length function -------------------------------------------------------------------

read_length <- function(file, ws, member, species, year) {

  # fleet      <- gsub("LEN ","", ws)
  fleet      <- stringr::word(gsub("FLEET ", "F", toupper(ws)), 2)
  lengthtype <- as.character(read_excel(file, sheet = ws, col_names = FALSE, col_types = "text", range = "A8", .name_repair = "minimal"))
  lengthunit <- as.character(read_excel(file, sheet = ws, col_names = FALSE, col_types = "text", range = "A9", .name_repair = "minimal"))
  unit       <- as.character(read_excel(file, sheet = ws, col_names = FALSE, col_types = "text", range = "F5", .name_repair = "minimal"))

  t <-
    suppressMessages(read_excel(file, sheet = ws, col_names = FALSE, na="NA", col_types = "text", range = "A10:E109", .name_repair = "minimal")) %>%
    setNames(c("length","quarter1","quarter2","quarter3","quarter4")) %>%
    gather(key=quarter, value=value, quarter1:quarter4) %>%
    mutate(member=member, species=species, year=year, file=basename(file), fleet=fleet, unit=unit,
           lengthtype=lengthtype, lengthunit = lengthunit)  %>%
    mutate(quarter = as.numeric(stringr::str_match_all(quarter, "[0-9]+")) ) %>%
    mutate(value   = as.numeric(gsub("\\s+", "", value) )) %>%
    mutate(length  = as.numeric(length)) %>%
    mutate(unit = gsub("[[:punct:]]", "", tolower(unit))) %>%
    mutate(lengthunit = gsub("[[:punct:]]", "", tolower(lengthunit))) %>%
    filter(!is.na(value)) %>%

    mutate(value = ifelse(unit == "individuals", value/1000, value),
           unit  = ifelse(unit == "individuals", "thousands", unit)) %>%
    mutate(value = ifelse(unit == "millions", value*1000, value),
           unit  = ifelse(unit == "millions", "thousands", unit)) %>%

    group_by(member, fleet, species, year, quarter, lengthtype) %>%
    arrange(member, fleet, species, year, quarter, lengthtype, length) %>%
    mutate(prop = value / sum(value)) %>%
    mutate(cumvalue = cumsum(value),
           cumprop = cumsum(prop)) %>%
    ungroup()


  # generate dataframe is not all empty
  if (nrow(t) > 0) {
    print(ws)
    return(t)
  }

} # end of function


# read_alk function -------------------------------------------------------------------

# file       <- file.list[1]
# worksheets <- excel_sheets(file)
# ws         <- worksheets[grepl("ALK", worksheets)][1]
# member     <- as.character(read_excel(file, sheet = 1, col_names = FALSE, col_types = "text", range = "F6"))
# species    <- as.character(read_excel(file, sheet = 1, col_names = FALSE, col_types = "text", range = "F5"))
# year       <- as.integer  (read_excel(file, sheet = 1, col_names = FALSE, col_types = "numeric", range = "K3"))
# i <- 3
# ws = "ALK F2"
# file = file.list[i]

read_alk <- function(file, ws, member, species, year) {

  # fleet      <- gsub("ALK ","", ws)
  fleet      <- stringr::word(gsub("FLEET ", "F", toupper(ws)), 2)
  lengthtype <- as.character(read_excel(file, sheet = ws, col_names = FALSE, col_types = "text", range = "A10", .name_repair = "minimal") )
  lengthunit <- as.character(read_excel(file, sheet = ws, col_names = FALSE, col_types = "text", range = "A11", .name_repair = "minimal") )

  # ages
  a <-
    suppressMessages(read_excel(file, sheet = ws, col_names = FALSE, col_types = "text", range = "B11:N11"))  %>%
    mutate_all( ~ paste0("age",.)) %>%
    as.character()

  t1 <-
    suppressMessages(read_excel(file, sheet = ws, col_names = FALSE, na="NA", col_types = "text", range = "A12:N111", .name_repair = "minimal")) %>%
    setNames(c("length",all_of(a))) %>%
    mutate(quarter=1) %>%
    gather(key=age, value=value, all_of(a)) %>%
    filter(!is.na(value)) %>%
    mutate(value   = as.numeric(gsub("\\s+", "", value) ))

  t2 <-
    suppressMessages(read_excel(file, sheet = ws, col_names = FALSE, na="NA", col_types = "text", range = "P12:AC111", .name_repair = "minimal")) %>%
    setNames(c("length",all_of(a))) %>%
    mutate(quarter=2) %>%
    gather(key=age, value=value, all_of(a)) %>%
    filter(!is.na(value)) %>%
    mutate(value   = as.numeric(gsub("\\s+", "", value) ))

  t3 <-
    suppressMessages(read_excel(file, sheet = ws, col_names = FALSE, na="NA", col_types = "text", range = "AE12:AR111", .name_repair = "minimal")) %>%
    setNames(c("length",all_of(a))) %>%
    mutate(quarter=3) %>%
    gather(key=age, value=value, all_of(a)) %>%
    filter(!is.na(value)) %>%
    mutate(value   = as.numeric(gsub("\\s+", "", value) ))

  t4 <-
    suppressMessages(read_excel(file, sheet = ws, col_names = FALSE, na="NA", col_types = "text", range = "AT12:BG111", .name_repair = "minimal")) %>%
    setNames(c("length",all_of(a))) %>%
    mutate(quarter=4) %>%
    gather(key=age, value=value, all_of(a)) %>%
    filter(!is.na(value)) %>%
    mutate(value   = as.numeric(gsub("\\s+", "", value) ))

  t <-
    bind_rows(t1, t2, t3, t4) %>%
    mutate(member=member, species=species, year=year, file=basename(file), fleet=fleet,
           lengthtype=lengthtype, lengthunit = lengthunit)  %>%
    mutate(value   = as.numeric(gsub("\\s+", "", value) )) %>%
    mutate(length  = as.numeric(length)) %>%
    mutate(age     = as.integer(gsub("age","",age))) %>%
    mutate(lengthunit = gsub("[[:punct:]]", "", tolower(lengthunit))) %>%
    group_by(member, fleet, species, year, quarter, length) %>%
    mutate(prop = value / sum(value)) %>%
    ungroup()


  # generate dataframe is not all empty
  if (nrow(t) > 0) {

    print(ws)
    return(t)
  }

} # end of function


# read_canum function -------------------------------------------------------------------

# ws <- "CN F1"

read_canum <- function(file, ws, member, species, year) {

  # fleet      <- gsub("CN ","", ws)
  fleet      <- stringr::word(gsub("FLEET ", "F", toupper(ws)), 2)
  unit       <- as.character(read_excel(file, sheet = ws, col_names = FALSE, col_types = "text", range = "C11:N11", .name_repair = "minimal") )

  t <-
    suppressMessages(read_excel(file, sheet = ws, col_names = FALSE, na="NA", col_types = "text", range = "B12:N24", .name_repair = "minimal")) %>%
    setNames(c("age","canum_1","meanlength_1","meanweight_1","canum_2","meanlength_2","meanweight_2",
                     "canum_3","meanlength_3","meanweight_3","canum_4","meanlength_4","meanweight_4")) %>%
    mutate(age = as.integer(gsub("\\+","",age))) %>%
    pivot_longer(names_to = "tosplit", values_to="value", canum_1:meanweight_4) %>%
    mutate(unit=rep(unit,13)) %>%
    separate(tosplit, into=c("variable","quarter"), sep="_") %>%
    filter(!is.na(value)) %>%
    mutate(unit = gsub("[[:punct:]]", "", tolower(unit))) %>%
    mutate(member=member, species=species, year=year, file=basename(file), fleet=fleet)  %>%
    mutate(value   = as.numeric(gsub("\\s+", "", value) )) %>%

    mutate(value = ifelse(unit == "individuals", value/1000, value),
           unit  = ifelse(unit == "individuals", "thousands", unit)) %>%
    mutate(value = ifelse(unit == "millions", value*1000, value),
           unit  = ifelse(unit == "millions", "thousands", unit)) %>%

    arrange(variable, quarter, age) %>%
    ungroup()


  # generate dataframe is not all empty
  if (nrow(t) > 0) {

    print(ws)
    return(t)
  }

} # end of function



# i    <- 1
# file <- file.list[i]
# ws   <- "catch data"

# Create file list for import
file.list <- as.character(NA)
for (d in dir.list) {
  file.list   <- c(file.list,
                   list.files(path = d,recursive  = FALSE,pattern = "xls",full.names = TRUE,ignore.case= FALSE) )
}
file.list <- file.list[grepl("ALK", toupper(basename(file.list)))]
file.list <- file.list[!grepl("\\~", file.list)]
file.list <- file.list[!is.na(file.list)]

# create empty dataframes
catch <- sampling <- length <- alk <- alk2 <- canum <- data.frame(stringsAsFactors = FALSE)

i <- 8

# start file reading loop
for (i in 1:length(file.list)){

  # get name and number of worksheets in the file
  file        <- file.list[i]
  worksheets  <- readxl::excel_sheets(file)
  nworksheets <- length(worksheets)

  print(paste(i, basename(file), sep=" "))

  # generic information
  member <- as.character(read_excel(file, sheet = "START HERE", col_names = FALSE, col_types = "text", range = "F6", .name_repair = "minimal"))
  member <- toupper(member)
  species<- as.character(read_excel(file, sheet = "START HERE", col_names = FALSE, col_types = "text", range = "F5", .name_repair = "minimal"))
  year   <- as.integer  (read_excel(file, sheet = "START HERE", col_names = FALSE, col_types = "numeric", range = "K3", .name_repair = "minimal"))

  # catch
  for (ws in worksheets[grepl("catch", worksheets)]) {
    catch <- bind_rows(catch, read_catch(file=file, ws=ws, member=member, species=species, year=year)) }

  # sampling
  for (ws in worksheets[grepl("sampling", worksheets)]) {
    sampling <- bind_rows(sampling, read_sampling(file=file, ws=ws, member=member, species=species, year=year)) }

  # length
  for (ws in worksheets[grepl("LEN", worksheets)]) {
    length <- bind_rows(length, read_length(file=file, ws=ws, member=member, species=species, year=year)) }

  # alk
  for (ws in worksheets[grepl("ALK", worksheets)]) {
    alk <- bind_rows(alk, read_alk(file=file, ws=ws, member=member, species=species, year=year)) }

  # canum
  for (ws in worksheets[grepl("CN|CANUM", worksheets)]) {
    canum <- bind_rows(canum, read_canum(file=file, ws=ws, member=member, species=species, year=year)) }

} # end of i loop

# summarise to avoid any splits in files (e.g. Russia 2019)
catch    <- catch %>% group_by(variable, unit, quarter, year, fleet, member, species) %>% summarise(value = sum(value, na.rm=TRUE)) %>% ungroup()

sampling <- sampling %>% group_by(variable, unit, quarter, year, fleet, member, species) %>% summarise(value = sum(value, na.rm=TRUE)) %>% ungroup()

length   <- 
  length %>% 
  mutate(lengthtype = gsub(" ","", tolower(lengthtype))) %>% 
  group_by(unit, quarter, year, fleet, member, species, lengthtype, lengthunit, length) %>% 
  summarise(value = sum(value, na.rm=TRUE)) %>% 
  group_by(member, fleet, species, year, quarter, lengthtype) %>%
  arrange(member, fleet, species, year, quarter, lengthtype, length) %>%
  mutate(prop = value / sum(value)) %>%
  mutate(cumvalue = cumsum(value),
         cumprop = cumsum(prop)) %>%
  ungroup()

# canum
t1 <- canum %>% filter(variable %in% c("canum")) %>% 
  group_by(variable, unit, quarter, year, fleet, member, species, age) %>% 
  summarise(value = mean(value, na.rm=TRUE))
t2 <- canum %>% filter(variable %in% c("meanlength","meanweight")) %>% 
  group_by(variable, unit, quarter, year, fleet, member, species, age) %>% 
  summarise(value = mean(value, na.rm=TRUE))
canum <- bind_rows(t1,t2)

alk <- 
  bind_rows(alk, alk2) %>%
  mutate(lengthtype = gsub(" ","", tolower(lengthtype))) %>% 
  mutate(member = gsub("CHILE","Chile", member)) %>% 
  group_by(lengthtype, lengthunit, quarter, year, fleet, member, species, age, length) %>% 
  summarise(value = sum(value, na.rm=TRUE)) %>% 
  group_by(lengthtype, lengthunit, member, fleet, species, year, quarter, length) %>%
  mutate(prop = value / sum(value)) %>%
  ungroup()
  
# save(catch, sampling, length, alk, canum, file=file.path("D:/TEMP",fn))
# save(catch, sampling, length, alk, canum, file=file.path(data_path,fn))

```

SCW11 Jack Mackerel Data workshop

**Catch and sampling overviews**

Corresponding author: mpastoors@pelagicfish.eu

`r format(Sys.time(), '%d/%m/%Y')`

Files by year

```{r echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

t <-
  data.frame(file=file.list, stringsAsFactors = FALSE) %>% 
  mutate(
    year = parse_number(stringr::word(file, 7, sep="/")),
    fn = basename(file)
  )

t[-1] %>% 
  group_by(year) %>% 
  do(add_row(., .after=0)) %>%
  
  pander::pandoc.table(., 
                     style        = "simple",
                     split.tables = 200, 
                     split.cells  = c(rep(7,10)),
                     justify      = "left",
                     missing      =" ",
                     big.mark     = '', 
                     round        = 3)
```

##### page break

Overview of Jack mackerel catch and sampling overviews 2018-2020 generated from an automated template reader. The overview is based on the 'old' ageing method for the Chilean data. 

Catch by quarter

```{r echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

# plot catch
tc <-
  catch %>%
  mutate(value=as.numeric(value)) %>% 
  filter(value > 0) %>% 
  mutate(quarter = paste0("Q", quarter)) %>%
  mutate(variable = ifelse(variable %in% c("retained","discards"), "estimated_catch", variable)) %>% 
  mutate(variable = gsub(" ","_", variable)) %>% 
  
  # sum catch components
  group_by(year, quarter, fleet, member, variable, unit, species) %>% 
  summarise(value = sum(value, na.rm=TRUE)) %>% 
  
  # select one catch component for plotting; prefer estimated catch
  group_by(year, quarter, fleet, member, unit, species) %>% 
  mutate(n=n()) %>% 
  filter(!(n==2 & variable == "official landings")) %>% 
  dplyr::select(-n) %>% 
  ungroup()

tc %>%   
  ggplot(aes(x=member, y=value, group=year)) + 
  theme_publication() +
  theme(panel.spacing     = unit(2, "mm"),
        legend.key.width  = unit(1, "cm"),
        legend.title      = element_blank(),
        panel.grid.minor  = element_blank(),
        # legend.position = "null",
        text              = element_text(size=12) ) +
  
  labs(x = "quarter", y = "catch (tonnes)") +
  
  geom_bar(aes(fill=quarter), stat="identity", alpha=1) +
  scale_y_continuous(labels = comma) +
  facet_grid(year~fleet, scales="free_x", space="free_x")
# facet_grid(.~quarter)

```

##### page break

Number of length measurements

```{r echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

# plot sampled ages
sampling %>%
  filter(variable=="measured") %>% 
  mutate(value=as.numeric(value)) %>% 
  mutate(quarter = paste0("Q", quarter)) %>% 

  ggplot(aes(x=member, y=value)) + 
  theme_publication() +
  theme(panel.spacing     = unit(2, "mm"),
        legend.key.width  = unit(1, "cm"),
        legend.title      = element_blank(),
        panel.grid.minor  = element_blank(),
        # legend.position = "null",
        text              = element_text(size=12) ) +
  
  labs(x = "quarter", y = "measured") +
  
  geom_bar(aes(fill=quarter), stat="identity", alpha=1) +
  facet_grid(year~fleet, scales="free_x", space="free_x")
# facet_grid(.~quarter)

```

##### page break

Number of age measurements

```{r echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

# sampling %>% filter(fleet=="F1", member=="China") %>% View()

# plot sampled ages
sampling %>%
  filter(variable=="aged") %>% 
  mutate(value=as.numeric(value)) %>% 
  filter(value > 0) %>% 
  mutate(quarter = paste0("Q", quarter)) %>% 

  ggplot(aes(x=member, y=value)) + 
  theme_publication() +
  theme(panel.spacing     = unit(2, "mm"),
        legend.key.width  = unit(1, "cm"),
        legend.title      = element_blank(),
        panel.grid.minor  = element_blank(),
        # legend.position = "null",
        text              = element_text(size=12) ) +
  
  labs(x = "quarter", y = "aged") +
  
  geom_bar(aes(fill=quarter), stat="identity", alpha=1) +
  facet_grid(year~fleet, scales="free_x", space="free_x")

```

##### page break

Number of age measurements per ton

```{r echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

# sampling %>% filter(fleet=="F1", member=="China") %>% View()

# plot sampled ages
t1 <-
  sampling %>%
  dplyr::select(-unit) %>% 
  filter(variable=="aged") %>% 
  mutate(value=as.numeric(value)) %>% 
  filter(value > 0) %>% 
  mutate(quarter = paste0("Q", quarter)) %>% 
  group_by(year, quarter, fleet, member, species)  %>% 
  pivot_wider(names_from = "variable", values_from = "value") 

t2 <-
  tc %>% 
  dplyr::select(-unit) %>% 
  group_by(year, quarter, fleet, member, species)  %>% 
  pivot_wider(names_from = variable, values_from = value) 
  
full_join(t1, t2, by=c("year","quarter","fleet","member","species")) %>% 
  mutate(ages_1000ton = ifelse(is.na(estimated_catch), 
                               aged / (official_landings/1000), 
                               aged / (estimated_catch/1000) )) %>% 
  filter(!is.na(ages_1000ton)) %>% 
  # View()

  ggplot(aes(x=member, y=ages_1000ton)) + 
  theme_publication() +
  theme(panel.spacing     = unit(2, "mm"),
        legend.key.width  = unit(1, "cm"),
        legend.title      = element_blank(),
        panel.grid.minor  = element_blank(),
        # legend.position = "null",
        text              = element_text(size=12) ) +
  
  labs(x = "quarter", y = "n aged/1000 ton") +
  
  geom_bar(aes(fill=quarter), stat="identity", alpha=1, position=position_dodge2(width=0.9, preserve="single")) +
  facet_grid(year~fleet, scales="free_x", space="free_x")

```

##### page break

Cumulative length proportions 

```{r echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

# plot cumulative proportions at length
length %>%   
  filter(tolower(lengthtype) == "forklength") %>% 
  mutate(quarter = paste0("Q", quarter)) %>% 
  mutate(yearquarter = paste0(year, quarter)) %>% 
  
  ggplot(aes(x=length, y=cumprop, group=member)) + 
  theme_publication() +
  theme(panel.spacing     = unit(2, "mm"),
        legend.key.width  = unit(1, "cm"),
        legend.title      = element_blank(),
        panel.grid.minor  = element_blank(),
        # legend.position = "null",
        text              = element_text(size=12) ) +
  
  labs(x = "length", y = "prop") +
  
  geom_point(aes(colour=member, group=member), alpha=1, size=0.2) +
  geom_line(aes(colour=member, group=member), alpha=1) +
  guides(colour = guide_legend(nrow = 1)) +
  facet_grid(year+quarter~fleet)
  # facet_grid(fleet~yearquarter)
  # facet_grid(.~quarter)

```

##### page break

ALK

```{r echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

# plot age-length keys
alk %>%   
  filter(prop > 0) %>% 
  # filter(year == 2019) %>% 
  mutate(quarter = paste0("Q", quarter)) %>% 
  mutate(yearquarter = paste0(year, quarter)) %>% 
  
  ggplot(aes(x=length, y=age, group=quarter)) + 
  theme_publication() +
  theme(panel.spacing     = unit(2, "mm"),
        legend.key.width  = unit(1, "cm"),
        legend.title      = element_blank(),
        panel.grid.minor  = element_blank(),
        # legend.position = "null",
        text              = element_text(size=12) ) +
  
  labs(x = "length", y = "age") +
  
  geom_point(aes(size=prop, colour=member), alpha=0.5) +
  scale_size(range=c(0.1,2)) +
  guides(size=FALSE) +
  # facet_grid(fleet~quarter)
  facet_grid(year~quarter)

```

```{r echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

# plot age-length keys
alk %>%   
  filter(prop > 0) %>% 
  # filter(year == 2019) %>% 
  mutate(quarter = paste0("Q", quarter)) %>% 
  group_by(member, quarter, length, age) %>%
  summarize(value = sum(value, na.rm=TRUE)) %>% 
  group_by(member, quarter, length) %>% 
  mutate(prop = value / sum(value)) %>% 
  
  ggplot(aes(x=length, y=age, group=quarter)) + 
  theme_publication() +
  theme(panel.spacing     = unit(2, "mm"),
        legend.key.width  = unit(1, "cm"),
        legend.title      = element_blank(),
        panel.grid.minor  = element_blank(),
        # legend.position = "null",
        text              = element_text(size=12) ) +
  
  # labs(x = "length", y = "age") +
  
  geom_point(aes(size=prop, colour=member), alpha=0.5) +
  scale_size(range=c(0.1,2)) +
  guides(size=FALSE) +
  scale_y_continuous(breaks=seq(1,12,1)) +
  # facet_grid(fleet~quarter)
  facet_grid(quarter~member)

```

```{r echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

# plot age-length keys
alk %>%   
  filter(prop > 0) %>% 
  filter(length == 30) %>% 
  # filter(year == 2019) %>% 
  mutate(quarter = paste0("Q", quarter)) %>% 
  group_by(member, quarter, length, age) %>%
  summarize(value = sum(value, na.rm=TRUE)) %>% 
  group_by(member, quarter, length) %>% 
  mutate(prop = value / sum(value)) %>% 
  
  ggplot(aes(x=age, y=prop, group=member)) + 
  theme_publication() +
  theme(panel.spacing     = unit(2, "mm"),
        legend.key.width  = unit(1, "cm"),
        legend.title      = element_blank(),
        panel.grid.minor  = element_blank(),
        # legend.position = "null",
        text              = element_text(size=12) ) +
  
  # labs(x = "length", y = "age") +
  
  geom_point(aes(colour=member)) +
  geom_line(aes(colour=member)) +
  scale_x_continuous(breaks=seq(1,12,1)) +
  # facet_grid(fleet~quarter)
  facet_wrap(~quarter)

```

##### page break

CANUM

```{r echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

# plot canum
canum %>%   
  # filter(year == myyear) %>%
  filter(variable == "canum") %>% 
  mutate(quarter = paste0("Q", quarter)) %>% 
  mutate(yearquarter = paste0(year, quarter)) %>% 
  group_by(year, fleet, member, age) %>% 
  summarize(value = sum(value, na.rm=TRUE)) %>% 
  
  ggplot(aes(x=age, y=value, group=member)) + 
  theme_publication() +
  theme(panel.spacing     = unit(2, "mm"),
        legend.key.width  = unit(1, "cm"),
        legend.title      = element_blank(),
        panel.grid.minor  = element_blank(),
        # legend.position = "null",
        text              = element_text(size=12) ) +
  
  labs(x = "age", y = "canum") +
  
  geom_bar(aes(fill=member, group=member), stat="identity", alpha=1) +
  facet_grid(fleet~year)
# facet_grid(.~quarter)





```



