---
output: 
  word_document:
    reference_docx:  ../PFA_report_template_v1.4.dotx
---

```{r setup, include=FALSE}

# ---------------------------------------------------------------------------------------
# Offshore CPUE analysis v2019 compare SPRFMO and EU data
#
# 17/09/2019 Benchmark type analysis for SPRFMO 2019; includes China in CPUE series; uses only data from SPRFM secretariat
# ---------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

rm(list=ls())

options(max.print=1000000)

# Libraries
library(rmarkdown)     # rmarkdown functionality
library(pander)        # tables
library(lubridate)     # data handling
library(reshape2)      # reshaping data; e.g. cast
library(readxl)        # excel reader
library(broom)         # clean up statistics
library(scales)        # pretty scales
library(stringr)       # string manipulations
library(tidyverse)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(mgcv)          # tensor spline for gam
library(lme4)
library(MASS)
library(captioner)    # captioning of figures and tables

# default settings for tables
panderOptions('table.alignment.default', function(df) ifelse(sapply(df, is.numeric), 'right', 'left'))

# To number figures and tables
fig_nums <- captioner::captioner(prefix = "Figure ")
tab_nums <- captioner::captioner(prefix = "Table ")

# set onedrive directory
onedrive <- file.path(Sys.getenv('USERPROFILE'), 'PFA/PFA team site - PRF') 

# load spatial data
load(file.path(onedrive,"rdata/world.df.RData"))
load(file.path(onedrive,"rdata/fao.df.RData"))
# load(file.path(onedrive,"rdata/eez.df.RData"))
# load(file.path(onedrive,"rdata/fao.RData"))
# load(file.path(onedrive,"rdata/eez.RData"))

# Source my personal utilities
source("../../prf/r/my utils.R")
source("../../gisland/r/geo_inside.R")

data_path <- "D:/SPRFMO/data"

# year settings
fy <- 2006
ly <- 2018
ay <- 2019

# create offshore_all object with environmental data
offshore_eu <-
  get(load(file.path(data_path, "offshore_all.RData"))) %>% 
  filter(vesselcp=="EU") %>% 
  mutate(source="eu")

offshore_eu_sprfmo <-
  get(load(file.path(data_path, "offshore_all_sprfmo.RData"))) %>% 
  filter(vesselcp=="EU") %>% 
  mutate(source="sprfmo")

offshore_eu_comb <-
  bind_rows(offshore_eu, offshore_eu_sprfmo)


```

**Compare EU data in SPRFMO database and from EU data itself**

M.A. Pastoors 

mpastoors@pelagicfish.eu

`r format(Sys.time(), '%d/%m/%Y')`


**Number of vessels participating in the fishery**

```{r echo=FALSE,  message=FALSE, warning=FALSE}

# number of vessels

offshore_eu_comb %>%
  filter(species == "CJM") %>% 
  filter(year %in% 2009:2012) %>% 
  group_by(source, vesselcode2, vesselname, year) %>% 
  filter(row_number() == 1) %>% 
  arrange(year, source, vesselname) %>% 
  View()


offshore_eu_comb %>%
  filter(species == "CJM") %>% 
  group_by(source, vesselcode2, year) %>% 
  filter(row_number() == 1) %>% 
  group_by(source, year) %>% 
  summarize(nvessels = n()) %>% 
  
  mutate(year = as.character(year)) %>% 
  dcast(year ~ source, value.var="nvessels", sum) %>% 
  pandoc.table(., 
               style        = "simple",
               split.tables = 100, 
               split.cells  = c(rep(7,10)),
               justify      = "right",
               missing      =".",
               big.mark     = ',', 
               round        = c(0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0))

```

*`r tab_nums("nvessels")`*

##### page break

**Total catch of jack mackerel per year**

```{r echo=FALSE,  message=FALSE, warning=FALSE}

tab_nums(name    = "cjmcatch", 
         caption = "Total catch of Jack mackerel by contracting party",
         display = FALSE)


# summed catch of CJM
offshore_eu_comb %>% 
  filter(species == "CJM") %>% 
  
  group_by(source, year, species) %>% 
  summarize(catch = sum(catch, na.rm=TRUE)) %>% 
  
  ungroup() %>% 
  
  mutate(year = as.character(year)) %>% 
  dcast(year ~ source, value.var="catch", sum, margins=c("source", "year")) %>% 
  pandoc.table(., 
               style        = "simple",
               split.tables = 100, 
               split.cells  = c(rep(7,10)),
               justify      = "right",
               missing      =".",
               big.mark     = ',', 
               round        = c(0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0))

```

*`r tab_nums("cjmcatch")`*

##### page break

**Summed fishing hours (when available)**

I.e. numbers of hours fished. The data for 2017 are incomplete. 

```{r echo=FALSE,  message=FALSE, warning=FALSE}

tab_nums(name    = "haulduration", 
         caption = "Summed fishing hours by contracting party",
         display = FALSE)

# Summed haul duration
offshore_eu_comb %>% 
  group_by(source, vesselname, shootdatetime, shootlat, shootlon) %>% 
  filter(row_number() == 1) %>% 
  
  group_by(source, year) %>% 
  summarize(duration = as.integer(sum(duration, na.rm=TRUE))) %>% 
  mutate(year = as.character(year)) %>% 
  dcast(year ~ source, value.var="duration", sum, margins=c("source", "year")) %>% 
  pandoc.table(., 
               style        = "simple",
               split.tables = 100, 
               split.cells  = c(rep(7,10)),
               justify      = "right",
               missing      =".",
               big.mark     = ',', 
               round        = c(0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0))

```

*`r tab_nums("haulduration")`*

##### page break

**Number of fishing days (defined as days when a haul has been reported)**

```{r echo=FALSE,  message=FALSE, warning=FALSE}

tab_nums(name    = "fishingdays", 
         caption = "Number of fishing days by contracting party",
         display = FALSE)

# Number of fishing days

offshore_eu_comb %>% 
  group_by(source, vesselname, shootdatetime, shootlat, shootlon) %>% 
  filter(row_number() == 1) %>% 
  
  # fishing days by vessel
  group_by(source, vesselname, year) %>% 
  summarize(fishingdays = n_distinct(day)) %>% 
  
  # fishing days by CP
  group_by(source, year) %>% 
  summarize(fishingdays = sum(fishingdays, na.rm=TRUE)) %>% 
  
  mutate(year = as.character(year)) %>% 
  dcast(year ~ source, value.var="fishingdays", sum, margins=c("source", "year")) %>% 
  pandoc.table(., 
               style        = "simple",
               split.tables = 100, 
               split.cells  = c(rep(7,10)),
               justify      = "right",
               missing      =".",
               big.mark     = ',', 
               round        = c(0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0))

```

*`r tab_nums("fishingdays")`*

**Number of hauls **

```{r echo=FALSE,  message=FALSE, warning=FALSE}

tab_nums(name    = "nhauls", 
         caption = "Number of hauls by contracting party",
         display = FALSE)

# number of hauls
offshore_eu_comb %>% 
  group_by(source, vesselname, shootdatetime, shootlat, shootlon) %>% 
  filter(row_number() == 1) %>% 
  
  group_by(source, year) %>% 
  summarize(nhauls = n()) %>% 
  mutate(year = as.character(year)) %>% 
  dcast(year ~ source, value.var="nhauls", sum, margins=c("source", "year")) %>% 
  pandoc.table(., 
               style        = "simple",
               split.tables = 100, 
               split.cells  = c(rep(7,10)),
               justify      = "right",
               missing      =".",
               big.mark     = ',', 
               round        = c(0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0))

```

*`r tab_nums("nhauls")`*

##### page break

**Length of the fishing season (defined as the number of days between the first haul and the last haul in a year)**

```{r echo=FALSE,  message=FALSE, warning=FALSE}

tab_nums(name    = "fishingseason", 
         caption = "Length of the fishing season (days) by Contracting Party",
         display = FALSE)

# length of fishing season

offshore_eu_comb %>% 
  filter(species == "CJM") %>% 
  group_by(source, vesselcode2, shootdatetime, shootlat, shootlon) %>% 
  filter(row_number() == 1) %>% 
  
  group_by(source, year) %>% 
  summarize(ndays = max(day, na.rm=TRUE) - min(day, na.rm=TRUE) +1) %>% 
  mutate(year = as.character(year)) %>% 
  dcast(year ~ source, value.var="ndays", mean, margins=c("source", "year")) %>% 
  pandoc.table(., 
               style        = "simple",
               split.tables = 100, 
               split.cells  = c(rep(7,10)),
               justify      = "right",
               missing      =".",
               big.mark     = ',', 
               round        = c(0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0))

```

*`r tab_nums("fishingseason")`*



##### page break

**EU Haul positions by source and year**

The yearly postions of Jack mackerel fishery of the EU fleets. 

```{r echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

invisible(gc())

offshore_eu_comb %>% 
  # offshore_all %>% 
  # filter(species == "CJM") %>% 
  filter(!is.na(shootlon) | !is.na(shootlat)) %>% 
  filter(!is.na(year)) %>% 

  ggplot(aes(x=shootlon, y=shootlat)) +
  theme_publication() +
  theme(axis.title       = element_blank(), 
        text             = element_text(size=12),
        legend.key.width = unit(1, "cm"),
        panel.spacing    = unit(0.1, "lines") ) +  
  
  coord_quickmap(xlim=c(-120,-50) , ylim=c(-50,-10)) +
  geom_polygon(data=world.df, aes(long,lat,group=group), 
               fill="cornsilk", size=0.25,color="gray15", alpha=0.7) +
  
  geom_point(aes(colour=source), size=0.8, alpha=0.5) +
  
  ggtitle("EU haul positions by source and year") +
  facet_wrap(~year, ncol=3)
  # facet_grid(rows=vars(vesselcp), cols= vars(year), drop=FALSE)

```

*Figure 2.3. Haul positions by contracting party. Colours indicate the different contracting parties participating.*

```{r echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

invisible(gc())

offshore_eu_comb %>% 
  filter(!is.na(shootlon) | !is.na(shootlat)) %>% 
  filter(!is.na(year)) %>% 
  filter(year == 2008) %>% 
  filter(species == "CJM") %>% 
  filter(tolower(vesselname) == "annelies ilena")  %>%
  mutate(catch = ifelse(source=="sprfmo", 1000*catch, catch)) %>% 
  ggplot(aes(x=shootdatetime, y=catch)) +
  theme_publication() +
  theme(axis.title       = element_blank(), 
        text             = element_text(size=12),
        legend.key.width = unit(1, "cm"),
        panel.spacing    = unit(0.1, "lines") ) +  
  
  geom_point(aes(colour=source), size=0.8, alpha=0.5) 
  # facet_grid(rows=vars(vesselcp), cols= vars(year), drop=FALSE)

```