---
output: 
  powerpoint_presentation:
    reference_doc: ../powerpoint_template.pptx
---

```{r setup, include=FALSE}

# ---------------------------------------------------------------------------------------
# Jack mackerel CPUE analysis for offshore fleet in SPRFMO area (EU, Russia, Vanuatu, Korea + China)
#
# 23/09/2017 First coding of the CPUE analysis
# 08/05/2018 Added the fleets of Russia, Vanuatu and Korea
# 23/05/2018 Added data quality checking
# 28/05/2018 Version presented at the SPRFMO benchmark
# 08/08/2018 Version that includes 2017 data
# 17/09/2019 Benchmark type analysis for SPRFMO 2019; includes China in CPUE series; 
#            uses only data from SPRFMO secretariat
# 19/09/2019 included comparison with old Chinese CPUE series
# 25/09/2019 final checking of code and results
# 05/10/2019 updated the HCI and El Nino data; previously that lead to curtailing the dataseries
# 07/10/2019 powerpoint version
# ---------------------------------------------------------------------------------------


require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

rm(list=ls())

options(max.print=1000000)

# Libraries
library(rmarkdown)     # rmarkdown functionality
library(pander)        # tables
library(lubridate)     # data handling
library(reshape2)      # reshaping data; e.g. cast
library(readxl)        # excel reader
library(broom)         # clean up statistics
library(scales)        # pretty scales
library(stringr)       # string manipulations
library(tidyverse)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(mgcv)          # tensor spline for gam
library(lme4)
library(MASS)
library(captioner)    # captioning of figures and tables
# library(ggplotify)     # capture graphics as grob/ggplots
library(mgcViz)         # GAM output to ggplot

# Source my personal utilities
source("../../prf/r/my utils.R")
source("../../gisland/r/geo_inside.R")

# default settings for tables
panderOptions('table.alignment.default', function(df) ifelse(sapply(df, is.numeric), 'right', 'left'))

# To number figures and tables
fig_nums <- captioner::captioner(prefix = "Figure ")
tab_nums <- captioner::captioner(prefix = "Table ")

# set paths
onedrive  <- file.path(Sys.getenv('USERPROFILE'), 'PFA/PFA team site - PRF') 
data_path <- "D:/SPRFMO/data"
dropbox   <- file.path(get_dropbox(),"jurel","archiv", "SC06_2018")

# load spatial data
load(file.path(onedrive,"rdata/world.df.RData"))
load(file.path(onedrive,"rdata/fao.df.RData"))
# load(file.path(onedrive,"rdata/eez.df.RData"))
# load(file.path(onedrive,"rdata/fao.RData"))
# load(file.path(onedrive,"rdata/eez.RData"))


# year settings
fy <- 2008
ly <- 2018
ay <- 2019

# Load old Chinese CPUE data
china <-
  read.csv(file=paste0(dropbox, "/Chinese CPUE 2001-2017.csv"), header=TRUE) %>% 
  lowcase() %>% 
  filter (year >= fy) 

# Load the El Nino data
elnino <- 
  read_excel(path=file.path(data_path, "elnino 20191005.xlsx"), col_names = TRUE) %>% 
  lowcase() %>% 
  gather(key=month, value=sst, m1:m12) %>% 
  mutate(month = as.numeric(gsub("m","", month))) %>% 
  mutate(ELE   = 0, 
         ELE   = ifelse(sst <= -0.5, -1, ELE),
         ELE   = ifelse(sst >= 0.5 ,  1, ELE)) %>% 
  arrange(year, month)

# Load the Humbold Current Index
hci <-
  read.csv(file=file.path(data_path,"HCI_v 20191005.csv"), skip=2, header=TRUE) %>% 
  lowcase()

# source the Offshore CPUE reader to read in all the CPUE data
# source("Offshore CPUE reader.r")

# create offshore_all object with environmental data
offshore_all <-
  # get(load(file.path(data_path, "offshore_all.RData"))) %>% 
  get(load(file.path(data_path, "offshore_all_sprfmo.RData"))) %>% 
  left_join(elnino, by=c("year","month")) %>% 
  left_join(hci, by=c("year","month")) %>% 
  mutate(shootlon = ifelse(shootlon == 0, NA, shootlon),
         shootlat = ifelse(shootlat == 0, NA, shootlat))

# create cjm by hour (for CPUE analysis)
cjm_byhour <-
  offshore_all %>% 
  filter(species  == "CJM") %>% 
  filter(year %in% fy:ly) %>% 
  
  # remove very small (artificial) durations
  filter(duration > 0.005) %>% 
  
  group_by (vesselcp, vesselcode2, shootdatetime, year, month, day, species) %>% 
  summarize(catch    = sum(catch, na.rm=TRUE),
            effort   = sum(duration, na.rm=TRUE),
            shootlat = mean(shootlat, na.rm=TRUE),
            shootlon = mean(shootlon, na.rm=TRUE),
            sst      = mean(sst, na.rm=TRUE),
            ELE      = mean(ELE, na.rm=TRUE),
            hci      = mean(hci, na.rm=TRUE)) %>% 
  mutate(catch        = round(catch),
         lcatch      = log(catch + 0.1),
         effort      = ifelse(!is.na(effort) & effort > 0, effort, NA),
         cpue        = ifelse(!is.na(effort) & effort > 0, catch/effort, NA),
         lcpue       = ifelse(!is.na(effort), log(catch/effort+0.1), NA)) %>% 
  ungroup() %>% 
  mutate_at(c("year","month", "vesselcp", "vesselcode2","ELE"), funs(as.factor)) %>% 
  drop_na(c("vesselcp","vesselcode2","year","month","catch","effort","cpue",
            "shootlat","shootlon","sst","ELE", "hci"))

# offshore_all %>% filter(vesselcp=="EU" & year == 2008) %>% count_not_na() # ggplot(aes(effort)) + geom_histogram()
# offshore_all %>% filter(vesselcp=="EU" & year == 2008) %>% filter(is.na(duration)) %>% View()
# cjm_byhour %>% ggplot(aes(effort)) + geom_histogram() + facet_wrap(~year)


# cjm by day (for CPUE analysis)
cjm_byday <-
  
  offshore_all %>% 
  filter(species  == "CJM") %>% 
  filter(year %in% fy:ly) %>% 
  
  group_by(vesselcp, vesselcode2, year, month, day, species) %>% 
  summarize(catch    = sum(catch, na.rm=TRUE),
            shootlat = mean(shootlat, na.rm=TRUE),
            shootlon = mean(shootlon, na.rm=TRUE),
            sst      = mean(sst, na.rm=TRUE),
            ELE      = mean(ELE, na.rm=TRUE),
            hci      = mean(hci, na.rm=TRUE)) %>% 
  mutate(catch        = round(catch),
         lcatch      = log(catch + 0.1),
         effort      = 1,
         cpue        = catch,
         lcpue       = log(catch + 0.1)) %>% 
  ungroup() %>% 
  mutate_at(c("year","month", "vesselcp", "vesselcode2","ELE"), funs(as.factor)) %>% 
  drop_na(c("vesselcp","vesselcode2","year","month","catch","effort","cpue",
            "shootlat","shootlon","sst","ELE", "hci"))

# cjm_byday %>% ggplot(aes(cpue)) + geom_histogram() + facet_wrap(~year)
# offshore_all %>% filter(catch==0, vesselcp=="EU") %>% group_by(vesselcp, year) %>% summarize(n=n()) %>% View()
# offshore_all %>% filter(catch==0, vesselcp=="EU") %>% group_by(vesselcp, year, vesselname) %>% summarize(n=n()) %>% View()
# offshore_all %>% filter(catch==0, vesselcp=="EU", species=="CJM") %>%  View()

# cjm by day (for CPUE analysis)
cjm_byweek <-
  offshore_all %>% 
  filter(species == "CJM") %>% 
  # filter(vesselcp != "KOR") %>% 
  filter(year %in% (fy:ly)) %>% 
  mutate(week        = week(shootdatetime)) %>% 
  group_by(vesselcp, vesselcode2, year, month, week, species) %>% 
  summarize(catch    = sum(catch, na.rm=TRUE),
            effort   = n_distinct(day),
            shootlat = mean(shootlat, na.rm=TRUE),
            shootlon = mean(shootlon, na.rm=TRUE),
            sst      = mean(sst, na.rm=TRUE),
            ELE      = mean(ELE, na.rm=TRUE),
            hci      = mean(hci, na.rm=TRUE)) %>% 
  mutate(catch        = round(catch),
         lcatch      = log(catch + 0.1),
         cpue        = catch / effort, 
         lcpue       = log(catch/effort + 0.1)) %>% 
  ungroup() %>% 
  filter(catch > 0) %>%            # if no catch during a week, this is suspicious ?? CHECK!
  mutate_at(c("year","month", "vesselcp", "vesselcode2","ELE"), funs(as.factor)) %>% 
  drop_na(c("vesselcp","vesselcode2","year","month","catch","effort","cpue",
            "shootlat","shootlon","sst","ELE", "hci"))


# Calculate means for the prediction variables
mean_month <-
  cjm_byday %>% 
  group_by(month) %>% 
  summarize(lcpue = mean(lcpue, na.rm=TRUE)) %>% 
  ungroup() %>% 
  mutate(lcpue2 = abs(lcpue - mean(lcpue, na.rm=TRUE))) %>% 
  arrange(lcpue2) %>% 
  filter(row_number() == 1)

mean_vessel <-
  cjm_byday %>% 
  group_by(vesselcode2) %>% 
  summarize(lcpue = mean(lcpue, na.rm=TRUE)) %>% 
  ungroup() %>% 
  mutate(lcpue2 = abs(lcpue - mean(lcpue, na.rm=TRUE))) %>% 
  arrange(lcpue2) %>% 
  filter(row_number() == 1)


```

# Jack mackerel CPUE standardization for the combined offshore fleet including China

M.A. Pastoors & N.T. Hintzen (European Union)

`r format(Sys.time(), '%d/%m/%Y')`

---

# Number of vessels participating in the fishery

```{r echo=FALSE,  message=FALSE, warning=FALSE}

tab_nums(name    = "nvessels", 
         caption = "Number of vessels participating in the Jack mackerel fishery by Contracting Party",
         display = FALSE)

# number of vessels

offshore_all %>%
  filter(species == "CJM") %>% 
  group_by(vesselcp, vesselcode2, year) %>% 
  filter(row_number() == 1) %>% 
  group_by(vesselcp, year) %>% 
  summarize(nvessels = n()) %>% 
  
  mutate(year = as.character(year)) %>% 
  dcast(year ~ vesselcp, value.var="nvessels", sum, margins="vesselcp") %>% 
  pandoc.table(., 
               style        = "simple",
               split.tables = 100, 
               split.cells  = c(rep(7,10)),
               justify      = "right",
               missing      =".",
               big.mark     = ',', 
               round        = c(0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0))

```

---

# Total catch of jack mackerel per year

```{r echo=FALSE,  message=FALSE, warning=FALSE}

tab_nums(name    = "cjmcatch", 
         caption = "Total catch of Jack mackerel by contracting party",
         display = FALSE)


# summed catch of CJM
offshore_all %>% 
  filter(species == "CJM") %>% 
  
  group_by(vesselcp, year, species) %>% 
  summarize(catch = sum(catch, na.rm=TRUE)) %>% 
  
  ungroup() %>% 
  
  mutate(year = as.character(year)) %>% 
  dcast(year ~ vesselcp, value.var="catch", sum, margins=c("vesselcp", "year")) %>% 
  pandoc.table(., 
               style        = "simple",
               split.tables = 100, 
               split.cells  = c(rep(7,10)),
               justify      = "right",
               missing      =".",
               big.mark     = ',', 
               round        = c(0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0))

```

---

# Number of hauls

```{r echo=FALSE,  message=FALSE, warning=FALSE}

tab_nums(name    = "nhauls", 
         caption = "Number of hauls by contracting party",
         display = FALSE)

# number of hauls
offshore_all %>% 
  group_by(vesselname, shootdatetime, shootlat, shootlon) %>% 
  filter(row_number() == 1) %>% 
  
  group_by(vesselcp, year) %>% 
  summarize(nhauls = n()) %>% 
  mutate(year = as.character(year)) %>% 
  dcast(year ~ vesselcp, value.var="nhauls", sum, margins=c("vesselcp", "year")) %>% 
  pandoc.table(., 
               style        = "simple",
               split.tables = 100, 
               split.cells  = c(rep(7,10)),
               justify      = "right",
               missing      =".",
               big.mark     = ',', 
               round        = c(0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0))

```

---

# Mean catch per day of jack mackerel

```{r echo=FALSE,  message=FALSE, warning=FALSE}

tab_nums(name    = "meancatchperday", 
         caption = "Mean catch per day of Jack Mackerel",
         display = FALSE)

# CPUE per day averaged by year

cjm_byday %>% 
# offshore_all %>%
  filter(species == "CJM") %>%

  group_by(vesselcp, vesselcode2, year, day, species) %>% 
  summarize(cpue = sum(catch, na.rm=TRUE)) %>% 
  
  group_by(vesselcp, year, species) %>% 
  summarize(cpue = mean(cpue, na.rm=TRUE)) %>% 
  ungroup() %>% 
  
  mutate(year = as.character(year)) %>% 
  dcast(year ~ vesselcp, value.var="cpue", mean, margins=c("vesselcp", "year")) %>% 
  pandoc.table(., 
               style        = "simple",
               split.tables = 100, 
               split.cells  = c(rep(7,10)),
               justify      = "right",
               missing      =".",
               big.mark     = ',', 
               round        = c(0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0))

# offshore_all %>%
#   filter(species == "CJM") %>% 
#   group_by(vesselcode, year, day) %>% 
#   summarize(catch = sum(catch, na.rm=TRUE)) %>% 
#   group_by(vesselcode) %>% 
#   summarize(catch = mean(catch)) %>% 
#   arrange(-catch) %>% 
#   View()

```

---

# Haul positions by year

```{r echo=FALSE, fig.asp=0.4, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(name    = "haulpositionsbyyear", 
         caption = "Haul positions where Jack mackerel has been caught (by year). Colours indicate the different contracting parties",
         display = FALSE)

invisible(gc())

cjm_byhour %>% 
  # offshore_all %>% 
  # filter(species == "CJM") %>% 
  filter(!is.na(shootlon) | !is.na(shootlat)) %>% 
  filter(!is.na(year)) %>% 

  ggplot(aes(x=shootlon, y=shootlat)) +
  theme_publication() +
  theme(axis.title       = element_blank(), 
        text             = element_text(size=12),
        legend.key.width = unit(1, "cm"),
        panel.spacing    = unit(0.1, "lines") ) +  
  
  coord_quickmap(xlim=c(-120,-50) , ylim=c(-50,-10)) +
  geom_polygon(data=world.df, aes(long,lat,group=group), 
               fill="cornsilk", size=0.25,color="gray15", alpha=0.7) +
  
  geom_point(aes(colour=vesselcp), size=0.8, alpha=0.5) +
  
  # ggtitle("CJM haul positions by year") +
  facet_wrap(~year, ncol=6)
  # facet_grid(rows=vars(vesselcp), cols= vars(year), drop=FALSE)

```

---

# Comparison of different CPUE metrics: by hour, by day and by week

```{r echo=FALSE, fig.asp=0.5, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "comparecpuemetrics", 
  caption = "Jack mackerel CPUE metrics by hour, by day and by week, scaled to the maximum of the time series.",
  display = FALSE)

# CPUE per day averaged by year
byhour <-
  cjm_byhour %>% 
  group_by(vesselcp, year) %>% 
  summarize(
    catch_byhour  = mean(catch, na.rm=TRUE),
    cpue_byhour   = mean(cpue, na.rm=TRUE),
    effort_byhour = mean(effort, na.rm=TRUE)) %>% 
  group_by(vesselcp) %>% 
  mutate(
    catch_byhour = catch_byhour/max(catch_byhour, na.rm=TRUE),
    cpue_byhour = cpue_byhour/max(cpue_byhour, na.rm=TRUE),
    effort_byhour = effort_byhour/max(effort_byhour, na.rm=TRUE))

byday <-
  cjm_byday %>% 
  group_by(vesselcp, year) %>% 
  summarize(
    catch_byday  = mean(catch, na.rm=TRUE),
    cpue_byday   = mean(cpue, na.rm=TRUE),
    effort_byday = mean(effort, na.rm=TRUE)) %>% 
  group_by(vesselcp) %>% 
  mutate(
    catch_byday = catch_byday/max(catch_byday, na.rm=TRUE),
    cpue_byday = cpue_byday/max(cpue_byday, na.rm=TRUE),
    effort_byday = effort_byday/max(effort_byday, na.rm=TRUE))

byweek <-
  cjm_byweek %>% 
  group_by(vesselcp, year) %>% 
  summarize(
    catch_byweek  = mean(catch, na.rm=TRUE),
    cpue_byweek   = mean(cpue, na.rm=TRUE),
    effort_byweek = mean(effort, na.rm=TRUE)) %>% 
  group_by(vesselcp) %>% 
  mutate(
    catch_byweek = catch_byweek/max(catch_byweek, na.rm=TRUE),
    cpue_byweek = cpue_byweek/max(cpue_byweek, na.rm=TRUE),
    effort_byweek = effort_byweek/max(effort_byweek, na.rm=TRUE))

my.print <- "cpue"

bind_rows(byhour, byday, byweek) %>% 
  gather(key=metric, value=value, catch_byhour:effort_byweek) %>% 
  
  filter(grepl(my.print, metric)) %>% 
  
  mutate(year = as.numeric(as.character(year))) %>% 

  ggplot(aes(x=year, y=value, group=metric)) +
  theme_publication() +
  theme(axis.title       = element_blank(), 
        text             = element_text(size=12),
        legend.key.width = unit(1, "cm"),
        legend.key.size  = unit(2, "cm"),
        panel.spacing    = unit(0.1, "lines") ) +  
  
  geom_point(aes(colour=metric), size=2, alpha=0.5) +
  geom_path(aes(colour=metric)) +
  
  # ggtitle(paste0(my.print, " metrics (scaled to maximum)")) +
  facet_wrap(~vesselcp, ncol=5, scales="free_y") +
  scale_x_continuous(breaks=seq(fy, ly, by = 5)) +
  expand_limits(y=0)


```

---

# Log CPUE by week and yearly average Log CPUE

```{r echo=FALSE, fig.asp=0.5, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "lcpuebyweek", 
  caption = "Jack mackerel log CPUE (log(catch / ndays)) by week.",
  display = FALSE)

invisible(gc())

m <-
  cjm_byweek %>% 
  group_by(vesselcp, year) %>%
  summarise(lcatch   = mean(lcpue, na.rm=TRUE), 
            week     = mean(week, na.rm=TRUE)) %>% 
  mutate(date = as.numeric(as.character(year)) + week/52) %>% 
  data.frame() 

cjm_byweek %>%
  mutate(date   = as.numeric(as.character(year)) + week/52,
         lcatch = lcpue) %>%

  # mutate(date   = as.numeric(as.character(year)) + week/52,
  #        lcatch = log(catch + 0.1)) %>%
  
  ggplot(aes(date, lcatch)) +
  theme_publication() +
  # theme(legend.position="none")+
  geom_jitter(aes(colour=vesselcp)) +
  geom_line(data=m, aes(date, lcatch), colour="gray20", linetype="dashed", size=1) +
  geom_point(data=m, aes(date, lcatch)) +
  scale_x_continuous(breaks=seq(fy, ly, by = 2)) +
  facet_wrap(~vesselcp, ncol=5)

```

---

# El Nino effect and Humbold_current index

```{r echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "elnino", 
  caption = "El Nino temperature anomaly (blue line) and ELE indicator (red line). Humboldt Current Index (green line)",
  display = FALSE)

invisible(gc())

elnino %>% 
  left_join(hci, by=c("year","month")) %>% 
  mutate(date = year + (month-1)/12) %>% 
  filter(year >= fy & year <= ly) %>% 
  
  ggplot(aes(date, sst) ) +
  theme_publication() +
  # theme(legend.position="none") +
  geom_point(alpha=0.5, colour="blue") +
  geom_line(colour="blue") +
  geom_line(aes(y=ELE), colour="red") +
  
  geom_point(aes(y=hci), alpha=0.5, colour="green") +
  geom_line(aes(y=hci), colour="green") +
  geom_smooth(aes(y=hci), colour="green", span=0.2, se=FALSE) +
  
  scale_x_continuous(breaks = seq(fy, ly, by = 5)) 


```

---

# Modelling approach

- GLM/GAM models on weekly catch with log(effort) as offset
- year effect (as factor) as the main explanatory variableO
- Other potential explanatory variables vessel, month, CP, lat*long, SST anomaly, El Nino effect, HCI
- using AIC criteria to identify best fitting second, third etc. variable. 

- A leave-one-out analysis
- Analysis for each CP separately.  

---

# AIC analysis to identify first linear effect

```{r echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "firstfactor", 
  caption = "Negative binomial GLM with best fitting first linear effect",
  display = FALSE)

#- Select the first linear effect next to the year effect

store <- list()
for(iVar in c("month","ELE","lonlat","vesselcp","vesselcode2","sst", "hci")){
  # print(iVar)
  form       <- formula(paste("catch ~ year + offset(log(effort)) +",iVar))
  if(iVar == "lonlat")
    form     <- formula(paste("catch ~ year + offset(log(effort)) + shootlon*shootlat"))
  # print(form)
  store[[iVar]] <- glm.nb(form,data=cjm_byweek)
}

AIC <- 
  as.data.frame((do.call(rbind,lapply(store,AIC))[,1])) %>% 
  bind_cols(var = rownames(.)) %>% 
  setNames(c("aic","var")) %>% 
  arrange((aic)) %>% 
  mutate(var = factor(var, var))  # this is the way to set the factor to the desired sequence

# plot AIC
ggplot(AIC, aes(x=var, y=aic)) + theme_publication() + geom_point(size=3)


```

---

# Diagnostics for the first linear effect: vessel

```{r echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

# plot diagnostics
par(mfrow=c(2,2)); gam.check(store[["vesselcode2"]])

```

---

# Further (sequential) linear effects

- Month
- Lat * Long
- Environmental variables (SST anomaly, El Nino, HCI)

---

# Exploring the environmental variables: ELE selected


```{r echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

#- Explore the el nino effects

store <- list()
for(iVar in c("ELE","sst", "hci")){
  form          <- formula(paste("catch ~ year + month + vesselcode2 + shootlon*shootlat +
                                  offset(log(effort)) +",iVar))
  store[[iVar]] <- glm.nb(form,data=cjm_byweek)
}

AIC <- (as.data.frame((do.call(rbind,lapply(store,AIC))[,1]))) %>% 
  bind_cols(var = rownames(.)) %>% 
  setNames(c("aic","var")) %>% 
  arrange((aic)) %>% 
  mutate(var = factor(var, var))  # this is the way to set the factor to the desired sequence



```

```{r echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

tab_nums(
  name    = "ELEfactor", 
  caption = "ANOVA results for negative binomial GLM including the El Nino Effect ELE",
  display = FALSE)

# print ANOVA table
anova(store[["ELE"]])

```

```{r echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

tab_nums(
  name    = "SSTfactor", 
  caption = "ANOVA results for negative binomial GLM including the Sea Surface Temperature (SST) anomaly",
  display = FALSE)

# print ANOVA tables
anova(store[["sst"]])
```

```{r echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

tab_nums(
  name    = "HCIfactor", 
  caption = "ANOVA results for negative binomial GLM including the Humboldt Current Index HCI",
  display = FALSE)

# print ANOVA tables
anova(store[["hci"]])

```


---

# Final GAM model: Catch ~ offset(log(effort)) + year + vessel + month + s(lat-lon) + ELE

```{r echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "finalmodeleffects", 
  caption = "Jack mackerel Final GAM model estimates for selected effects",
  display = FALSE)

formglm <- 
  formula(paste("catch ~ year + month + vesselcode2 + shootlon*shootlat + ELE + offset(log(effort))"))

final     <- formula(paste("catch ~ year + vesselcode2 + month + s(shootlon,shootlat) + ELE +
                           offset(log(effort))"))

newdat    <- expand.grid(
  year        = as.factor(fy:ly),
  month       = as.factor(3),
  shootlon    = quantile(cjm_byweek$shootlon,probs=c(0.5)),
  shootlat    = quantile(cjm_byweek$shootlat,probs=c(0.5)),
  ELE         = as.factor(0),
  vesselcode2 = names(rev(sort(colSums(table(cjm_byweek$year,cjm_byweek$vesselcode2)))))[1],
                         effort=7)

finalMod  <- gam(final, data=cjm_byweek,
                 family=negbin(glm.nb(formglm,data=cjm_byweek)$theta))

# plot the estimates of the different effects
plot.gam(finalMod, all.terms=T, page=1)

```

---

# Final GAM model: spatial smoothing

```{r echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

plot.gamViz(finalMod, select=1)

```

---

# Standardized CPUE for Jack mackerel

```{r echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "cpueindexplot", 
  caption = "GAM standardized offshore fleet CPUE for jack mackerel",
  display = FALSE)

oldcpue <- read.csv(
  file="D:/SPRFMO/data/Offshore fleet standardized CPUE 2006-2017.csv", 
  stringsAsFactors = FALSE) %>% 
  mutate(type="standardized")

# calculate the predicted values and confidence intervals
pred      <- predict(finalMod,newdat,se.fit=T,type="link")
upr       <- exp(pred$fit + (1.96 * pred$se.fit))
lwr       <- exp(pred$fit - (1.96 * pred$se.fit))

# create df
df        <- data.frame(
               cpue = exp(pred$fit),
               upr  = upr, 
               lwr  = lwr,
               year = fy:ly,
               type = "standardized",
               assessmentyear = ay,
               stringsAsFactors = FALSE) %>% 
  bind_rows(oldcpue)


ggplot(df, aes(x=year, group=assessmentyear)) +
  theme_publication() +
  geom_ribbon(aes(ymin=lwr, ymax=upr, fill=factor(assessmentyear)), alpha=0.2) +
  geom_line(aes(y=cpue, colour=factor(assessmentyear)), size=1) +
  expand_limits(y=0) +
  labs(y="cpue") +
  scale_x_continuous(breaks=pretty_breaks()) +
  facet_wrap(~type, scales = "free_y")


```

---

# Standardized CPUE output table

```{r echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

tab_nums(
  name    = "anovafinalmod", 
  caption = "ANOVA results with final model GAM",
  display = FALSE)

# print ANOVA tables
anova(finalMod)

# print(plot.gamViz(store[["formgamy"]]), pages=1)

```

---

# Leave one out analysis

```{r echo=FALSE, fig.asp=0.5, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "leaveoneout", 
  caption = "Jack mackerel leave-one-out analysis (leaving out one of the fleets)",
  display = FALSE)

cjm_byweek_noEU  <- filter(cjm_byweek, vesselcp != "EU" )
cjm_byweek_noKOR <- filter(cjm_byweek, vesselcp != "KOR" )
cjm_byweek_noVUT <- filter(cjm_byweek, vesselcp != "VUT" )
cjm_byweek_noRUS <- filter(cjm_byweek, vesselcp != "RUS" )
cjm_byweek_noCHN <- filter(cjm_byweek, vesselcp != "CHN" )

finalMod_noEU  <- gam(final,
                      data=cjm_byweek_noEU,
                      family=negbin(glm.nb(formglm,
                                           data=cjm_byweek_noEU)$theta))
finalMod_noKOR <- gam(final,
                      data=cjm_byweek_noKOR,
                      family=negbin(glm.nb(formglm,
                                           data=cjm_byweek_noKOR)$theta))
finalMod_noVUT <- gam(final,
                      data=cjm_byweek_noVUT,
                      family=negbin(glm.nb(formglm,
                                           data=cjm_byweek_noVUT)$theta))

finalMod_noRUS <- gam(final,
                      data=cjm_byweek_noRUS,
                      family=negbin(glm.nb(formglm,
                                           data=cjm_byweek_noRUS)$theta))

finalMod_noCHN <- gam(final,
                      data=cjm_byweek_noCHN,
                      family=negbin(glm.nb(formglm,
                                           data=cjm_byweek_noCHN)$theta))

newdat_noEU <- expand.grid(
  year        = as.factor(min(as.numeric(as.character(cjm_byweek_noEU$year))):
                          max(as.numeric(as.character(cjm_byweek_noEU$year)))),
  month       = as.factor(3),
  shootlon    = quantile(cjm_byweek_noEU$shootlon,probs=c(0.5)),
  shootlat    = quantile(cjm_byweek_noEU$shootlat,probs=c(0.5)),
  ELE         = as.factor(0),
  vesselcode2 = names(rev(sort(colSums(table(cjm_byweek_noEU$year,cjm_byweek_noEU$vesselcode2)))))[1],
                         effort=7)

newdat_noKOR    <- expand.grid(
  year        = as.factor(min(as.numeric(as.character(cjm_byweek_noKOR$year))):
                          max(as.numeric(as.character(cjm_byweek_noKOR$year)))),
  month       = as.factor(3),
  shootlon    = quantile(cjm_byweek_noKOR$shootlon,probs=c(0.5)),
  shootlat    = quantile(cjm_byweek_noKOR$shootlat,probs=c(0.5)),
  ELE         = as.factor(0),
  vesselcode2 = names(rev(sort(colSums(table(cjm_byweek_noKOR$year,cjm_byweek_noKOR$vesselcode2)))))[1],
                         effort=7)

newdat_noVUT    <- expand.grid(
  year        = as.factor(min(as.numeric(as.character(cjm_byweek_noVUT$year))):
                          max(as.numeric(as.character(cjm_byweek_noVUT$year)))),
  month       = as.factor(3),
  shootlon    = quantile(cjm_byweek_noVUT$shootlon,probs=c(0.5)),
  shootlat    = quantile(cjm_byweek_noVUT$shootlat,probs=c(0.5)),
  ELE         = as.factor(0),
  vesselcode2 = names(rev(sort(colSums(table(cjm_byweek_noVUT$year,cjm_byweek_noVUT$vesselcode2)))))[1],
                         effort=7)

newdat_noRUS    <- expand.grid(
  year        = as.factor(min(as.numeric(as.character(cjm_byweek_noRUS$year))):
                          max(as.numeric(as.character(cjm_byweek_noRUS$year)))),
  month       = as.factor(3),
  shootlon    = quantile(cjm_byweek_noRUS$shootlon,probs=c(0.5)),
  shootlat    = quantile(cjm_byweek_noRUS$shootlat,probs=c(0.5)),
  ELE         = as.factor(0),
  vesselcode2 = names(rev(sort(colSums(table(cjm_byweek_noRUS$year,cjm_byweek_noRUS$vesselcode2)))))[1],
                         effort=7)

newdat_noCHN    <- expand.grid(
  year        = as.factor(min(as.numeric(as.character(cjm_byweek_noCHN$year))):
                          max(as.numeric(as.character(cjm_byweek_noCHN$year)))),
  month       = as.factor(3),
  shootlon    = quantile(cjm_byweek_noCHN$shootlon,probs=c(0.5)),
  shootlat    = quantile(cjm_byweek_noCHN$shootlat,probs=c(0.5)),
  ELE         = as.factor(0),
  vesselcode2 = names(rev(sort(colSums(table(cjm_byweek_noCHN$year,cjm_byweek_noCHN$vesselcode2)))))[1],
                         effort=7)

# calculate the predicted values and confidence intervals
pred_noEU       <- predict(finalMod_noEU ,newdat_noEU,se.fit=T,type="link")
pred_noKOR      <- predict(finalMod_noKOR,newdat_noKOR,se.fit=T,type="link")
pred_noVUT      <- predict(finalMod_noVUT,newdat_noVUT,se.fit=T,type="link")
pred_noRUS      <- predict(finalMod_noRUS,newdat_noRUS,se.fit=T,type="link")
pred_noCHN      <- predict(finalMod_noCHN,newdat_noCHN,se.fit=T,type="link")

# create df
df    <- data.frame(
               cpue = exp(pred_noEU$fit),
               upr  = exp(pred_noEU$fit + (1.96 * pred_noEU$se.fit)), 
               lwr  = exp(pred_noEU$fit - (1.96 * pred_noEU$se.fit)),
               year = min(as.numeric(as.character(cjm_byweek_noEU$year))):
                      max(as.numeric(as.character(cjm_byweek_noEU$year))),
               type = "noEU",
               stringsAsFactors = FALSE
               ) %>% 
  bind_rows(., data.frame(
               cpue = exp(pred_noKOR$fit),
               upr  = exp(pred_noKOR$fit + (1.96 * pred_noKOR$se.fit)), 
               lwr  = exp(pred_noKOR$fit - (1.96 * pred_noKOR$se.fit)),
               year = min(as.numeric(as.character(cjm_byweek_noKOR$year))):
                      max(as.numeric(as.character(cjm_byweek_noKOR$year))),
               type = "noKOR",
               stringsAsFactors = FALSE
               ) ) %>% 
  bind_rows(., data.frame(
               cpue = exp(pred_noVUT$fit),
               upr  = exp(pred_noVUT$fit + (1.96 * pred_noVUT$se.fit)), 
               lwr  = exp(pred_noVUT$fit - (1.96 * pred_noVUT$se.fit)),
               year = min(as.numeric(as.character(cjm_byweek_noVUT$year))):
                      max(as.numeric(as.character(cjm_byweek_noVUT$year))),
               type = "noVUT",
               stringsAsFactors = FALSE
               ) ) %>% 
  bind_rows(., data.frame(
               cpue = exp(pred_noRUS$fit),
               upr  = exp(pred_noRUS$fit + (1.96 * pred_noRUS$se.fit)), 
               lwr  = exp(pred_noRUS$fit - (1.96 * pred_noRUS$se.fit)),
               year = min(as.numeric(as.character(cjm_byweek_noRUS$year))):
                      max(as.numeric(as.character(cjm_byweek_noRUS$year))),
               type = "noRUS",
               stringsAsFactors = FALSE
               ) ) %>% 
  bind_rows(., data.frame(
               cpue = exp(pred_noCHN$fit),
               upr  = exp(pred_noCHN$fit + (1.96 * pred_noCHN$se.fit)), 
               lwr  = exp(pred_noCHN$fit - (1.96 * pred_noCHN$se.fit)),
               year = min(as.numeric(as.character(cjm_byweek_noCHN$year))):
                      max(as.numeric(as.character(cjm_byweek_noCHN$year))),
               type = "noCHN",
               stringsAsFactors = FALSE
               ) ) %>% 
  mutate(year = as.integer(year))


ggplot(df, aes(year)) +
  theme_publication() +
  theme(legend.position  = "none") +
  geom_ribbon(aes(ymin=lwr, ymax=upr, fill=type), alpha=0.2) +
  geom_line(aes(y=cpue, colour=type), size=1) +
  expand_limits(y=0) +
  labs(y="cpue") +
  scale_x_continuous(breaks=pretty_breaks()) +
  facet_wrap(~type, scales = "free_y", ncol=5)

```

---

# Single fleet analyses

```{r echo=FALSE, fig.asp=0.5, fig.align="center", message=FALSE, warning=FALSE}

cjm_byweek_onlyEU  <- filter(cjm_byweek, vesselcp == "EU" )
cjm_byweek_onlyKOR <- filter(cjm_byweek, vesselcp == "KOR" )
cjm_byweek_onlyVUT <- filter(cjm_byweek, vesselcp == "VUT" )
# cjm_byweek_onlyRUS <- filter(cjm_byweek, vesselcp == "RUS" )
cjm_byweek_onlyCHN <- filter(cjm_byweek, vesselcp == "CHN" )

finalMod_onlyEU  <- gam(final,
                      data=cjm_byweek_onlyEU,
                      family=negbin(glm.nb(formglm,
                                           data=cjm_byweek_onlyEU)$theta))
finalMod_onlyKOR <- gam(final,
                      data=cjm_byweek_onlyKOR,
                      family=negbin(glm.nb(formglm,
                                           data=cjm_byweek_onlyKOR)$theta))
finalMod_onlyVUT <- gam(final,
                      data=cjm_byweek_onlyVUT,
                      family=negbin(glm.nb(formglm,
                                           data=cjm_byweek_onlyVUT)$theta))

# finalMod_onlyRUS <- gam(final,
#                       data=cjm_byweek_onlyRUS,
#                       family=negbin(glm.nb(formglm,
#                                            data=cjm_byweek_onlyRUS)$theta))

finalMod_onlyCHN <- gam(final,
                      data=cjm_byweek_onlyCHN,
                      family=negbin(glm.nb(formglm,
                                           data=cjm_byweek_onlyCHN)$theta))

newdat_onlyEU <- expand.grid(
  year        = as.factor(unique(cjm_byweek_onlyEU$year)),
  month       = as.factor(3),
  shootlon    = quantile(cjm_byweek_onlyEU$shootlon,probs=c(0.5)),
  shootlat    = quantile(cjm_byweek_onlyEU$shootlat,probs=c(0.5)),
  ELE         = as.factor(0),
  vesselcode2 = names(rev(sort(colSums(table(cjm_byweek_onlyEU$year,cjm_byweek_onlyEU$vesselcode2)))))[1],
                         effort=7)

newdat_onlyKOR    <- expand.grid(
  year        = as.factor(unique(cjm_byweek_onlyKOR$year)),
  month       = as.factor(3),
  shootlon    = quantile(cjm_byweek_onlyKOR$shootlon,probs=c(0.5)),
  shootlat    = quantile(cjm_byweek_onlyKOR$shootlat,probs=c(0.5)),
  ELE         = as.factor(0),
  vesselcode2 = names(rev(sort(colSums(table(cjm_byweek_onlyKOR$year,cjm_byweek_onlyKOR$vesselcode2)))))[1],
                         effort=7)

newdat_onlyVUT    <- expand.grid(
  year        = as.factor(unique(cjm_byweek_onlyVUT$year)),
  month       = as.factor(3),
  shootlon    = quantile(cjm_byweek_onlyVUT$shootlon,probs=c(0.5)),
  shootlat    = quantile(cjm_byweek_onlyVUT$shootlat,probs=c(0.5)),
  ELE         = as.factor(0),
  vesselcode2 = names(rev(sort(colSums(table(cjm_byweek_onlyVUT$year,cjm_byweek_onlyVUT$vesselcode2)))))[1],
                         effort=7)

# newdat_onlyRUS    <- expand.grid(
#   year        = as.factor(unique(cjm_byweek_onlyRUS$year)),
#   month       = as.factor(3),
#   shootlon    = quantile(cjm_byweek_onlyRUS$shootlon,probs=c(0.5)),
#   shootlat    = quantile(cjm_byweek_onlyRUS$shootlat,probs=c(0.5)),
#   ELE         = as.factor(0),
#   vesselcode2 = names(rev(sort(colSums(table(cjm_byweek_onlyRUS$year,cjm_byweek_onlyRUS$vesselcode2)))))[1],
#                          effort=7)

newdat_onlyCHN    <- expand.grid(
  year        = as.factor(unique(cjm_byweek_onlyCHN$year)),
  month       = as.factor(3),
  shootlon    = quantile(cjm_byweek_onlyCHN$shootlon,probs=c(0.5)),
  shootlat    = quantile(cjm_byweek_onlyCHN$shootlat,probs=c(0.5)),
  ELE         = as.factor(0),
  vesselcode2 = names(rev(sort(colSums(table(cjm_byweek_onlyCHN$year,cjm_byweek_onlyCHN$vesselcode2)))))[1],
                         effort=7)

# calculate the predicted values and confidence intervals
pred_onlyEU       <- predict(finalMod_onlyEU ,newdat_onlyEU,se.fit=T,type="link")
pred_onlyKOR      <- predict(finalMod_onlyKOR,newdat_onlyKOR,se.fit=T,type="link")
pred_onlyVUT      <- predict(finalMod_onlyVUT,newdat_onlyVUT,se.fit=T,type="link")
# pred_onlyRUS      <- predict(finalMod_onlyRUS,newdat_onlyRUS,se.fit=T,type="link")
pred_onlyCHN      <- predict(finalMod_onlyCHN,newdat_onlyCHN,se.fit=T,type="link")

# create df
df    <- data.frame(
               cpue = exp(pred_onlyEU$fit),
               upr  = exp(pred_onlyEU$fit + (1.96 * pred_onlyEU$se.fit)), 
               lwr  = exp(pred_onlyEU$fit - (1.96 * pred_onlyEU$se.fit)),
               year = unique(cjm_byweek_onlyEU$year),
               type = "onlyEU",
               stringsAsFactors = FALSE
               ) %>% 
  bind_rows(., data.frame(
               cpue = exp(pred_onlyKOR$fit),
               upr  = exp(pred_onlyKOR$fit + (1.96 * pred_onlyKOR$se.fit)), 
               lwr  = exp(pred_onlyKOR$fit - (1.96 * pred_onlyKOR$se.fit)),
               year = unique(cjm_byweek_onlyKOR$year),
               type = "onlyKOR",
               stringsAsFactors = FALSE
               ) ) %>% 
  bind_rows(., data.frame(
               cpue = exp(pred_onlyVUT$fit),
               upr  = exp(pred_onlyVUT$fit + (1.96 * pred_onlyVUT$se.fit)), 
               lwr  = exp(pred_onlyVUT$fit - (1.96 * pred_onlyVUT$se.fit)),
               year = unique(cjm_byweek_onlyVUT$year),
               type = "onlyVUT",
               stringsAsFactors = FALSE
               ) ) %>% 
  # bind_rows(., data.frame(
  #              cpue = exp(pred_onlyRUS$fit),
  #              upr  = exp(pred_onlyRUS$fit + (1.96 * pred_onlyRUS$se.fit)), 
  #              lwr  = exp(pred_onlyRUS$fit - (1.96 * pred_onlyRUS$se.fit)),
  #              year = unique(cjm_byweek_onlyRUS$year),
  #              type = "onlyRUS",
  #              stringsAsFactors = FALSE
  #              ) ) %>% 
  bind_rows(., data.frame(
               cpue = exp(pred_onlyCHN$fit),
               upr  = exp(pred_onlyCHN$fit + (1.96 * pred_onlyCHN$se.fit)), 
               lwr  = exp(pred_onlyCHN$fit - (1.96 * pred_onlyCHN$se.fit)),
               year = unique(cjm_byweek_onlyCHN$year),
               type = "onlyCHN",
               stringsAsFactors = FALSE
               ) ) %>% 
  mutate(year = as.numeric(as.character(year)))


ggplot(df, aes(year)) +
  theme_publication() +
  theme(legend.position  = "none") +
  geom_ribbon(aes(ymin=lwr, ymax=upr, fill=type), alpha=0.2) +
  geom_line(aes(y=cpue, colour=type), size=1) +
  geom_point(aes(y=cpue, colour=type), size=1.5) +
  expand_limits(y=0) +
  labs(y="cpue") +
  scale_x_continuous(breaks=pretty_breaks()) +
  facet_wrap(~type, scales = "free_y", ncol=5)



```

---

# Discussion and conclusions

- standardizing all the CPUE data from the offshore fleets (China, EU, Korea, vanuatu and Russia)
- Same approach as during benchmark 2018 (but now with China)
- Same explanatory factors
- Including leave-one-out analysis and single-fleet analysis

---

# Acknowledgements

We would like to acknowledge the permission granted by the delegations of China, Russia, Vanuatu and Korea to utilize their haul-by-haul data for the analysis of standardized CPUE of the offshore fleet fishing for Jack mackerel. 